---
title: "CIPJF"
author: "Qinyuan Xing"
date: "2022/4/23"
output:
  pdf_document:
    
    toc: yes
    toc_depth: '3'
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = "markup", message = F)
options(scientific=T, digits = 3) 
```

## 1. data import and clean
```{r}
# data 
heart_disease = data.frame(read.csv("E:\\2022 Spring\\R\\heart_2020_cleaned.csv"))
str(heart_disease)
heart_disease$HeartDisease = ifelse(heart_disease$HeartDisease=="Yes",1,0)
heart_disease$Smoking = as.factor(heart_disease$Smoking)
heart_disease$AlcoholDrinking= as.factor(heart_disease$AlcoholDrinking)
heart_disease$Stroke = as.factor(heart_disease$Stroke)
heart_disease$DiffWalking = as.factor(heart_disease$DiffWalking)
heart_disease$Sex = as.factor(heart_disease$Sex)
heart_disease$AgeCategory = as.factor(heart_disease$AgeCategory)
heart_disease$Race = as.factor(heart_disease$Race)
heart_disease$Diabetic = as.factor(heart_disease$Diabetic)
heart_disease$PhysicalActivity = as.factor(heart_disease$PhysicalActivity)
heart_disease$GenHealth = as.factor(heart_disease$GenHealth)
heart_disease$Asthma = as.factor(heart_disease$Asthma)
heart_disease$KidneyDisease = as.factor(heart_disease$KidneyDisease)
heart_disease$SkinCancer = as.factor(heart_disease$SkinCancer)
summary(heart_disease)
```
## 2. Treatment variable = AlcoholDrinking (Yes/No)

### 2.1 Sampling
```{r alcohol}
data0 = heart_disease[heart_disease$AlcoholDrinking=="No",]
data1 = heart_disease[heart_disease$AlcoholDrinking=="Yes",]
a = 0.02
set.seed(1111)
s0 = data0[sample(nrow(data0),size = floor(a*nrow(data0))),]
s1 = data1[sample(nrow(data1),size = floor(a*nrow(data1))),]
sdata = rbind(s0,s1)
summary(sdata$AlcoholDrinking)
```

### 2.2 Estimate Propensity Score 
```{r}
# estimate Propensity score
logmod = glm(AlcoholDrinking~.-HeartDisease,data = sdata,family = binomial(link = "logit"))
sdata$ps = predict(logmod,type = "response")
```

```{r}
library(Hmisc)
histbackback(split(sdata$ps,sdata$AlcoholDrinking),main="PS before matching",probability = T)
```

### 2.3 Linear regression before matching
```{r}
# Linear regression before matching
lmod_before = lm(HeartDisease~AlcoholDrinking+BMI+Smoking+Stroke+PhysicalHealth+MentalHealth+DiffWalking+Sex+AgeCategory+Race+Diabetic+PhysicalActivity+GenHealth+SleepTime+Asthma+KidneyDisease+SkinCancer,data = sdata)
summary(lmod_before)$coefficients[2,]
confint(lmod_before,parm = 2,0.95)
```
Before matching,in the linear regression, the coef and P-value of AlcoholDrinking show that the effect of alcohol drinking is significant and people who like drinking are less likely to have heart disease.

### 2.4 PSM
```{r}
# PSM
library(MatchIt)
match.res = matchit(AlcoholDrinking~BMI+Smoking+Stroke+PhysicalHealth+MentalHealth+DiffWalking+Sex+AgeCategory+Race+Diabetic+PhysicalActivity+GenHealth+SleepTime+Asthma+KidneyDisease+SkinCancer,data = sdata,method = "nearest")
matchsum = summary(match.res,standardize = T)$sum.matched
data.frame(round(matchsum[,1:3],3))
# absolute value of SMD (Standardized Mean Difference) <0.1 shows matching is OK for each variable.
matched_data = match.data(match.res)

histbackback(split(matched_data$ps,matched_data$AlcoholDrinking),main="PS after matching",probability = T)
plot(match.res,type = "hist")
match.matrix = data.frame(match.res$match.matrix)
T_index = match(row.names(match.matrix),row.names(matched_data))
C_index = match(match.matrix$match.res.match.matrix,row.names(matched_data))
Tgroup = matched_data[T_index,]
Cgroup = matched_data[C_index,]
m1.test=t.test(Tgroup$HeartDisease,Cgroup$HeartDisease,paired = T)
m1.test
```

The test show that after matching, the effect of alcoholdrinking(estimate ATT) is not significant.

### 2.5 Linear regression after matching
```{r}
# LM after matching
lmod_after = lm(HeartDisease~AlcoholDrinking+BMI+Smoking+Stroke+PhysicalHealth+MentalHealth+DiffWalking+Sex+AgeCategory+Race+Diabetic+PhysicalActivity+GenHealth+SleepTime+Asthma+KidneyDisease+SkinCancer,data = matched_data)
summary(lmod_after)$coefficients[2,]
confint(lmod_after,2,0.95)

```

### 2.6 Results
```{r}
#Results
alcohol.res = data.frame(method = c("LR before","PSM","LR after"),est_ATT=c(summary(lmod_before)$coefficients[2,1],m1.test$estimate,summary(lmod_after)$coefficients[2,1]),CI.lower = c(confint(lmod_before,2,0.95)[1],m1.test$conf.int[1],confint(lmod_after,2,0.95)[1]),CI.upper= c(confint(lmod_before,2,0.95)[2],m1.test$conf.int[2],confint(lmod_after,2,0.95)[2]),Pvalue = c(summary(lmod_before)$coefficients[2,4],m1.test$p.value,summary(lmod_after)$coefficients[2,4]),significant=c("Yes","No","No"))
alcohol.res
```

## 3. Treatment variable = Smoking(Yes/No)

### 3.1 Sampling
```{r Smoke}
data.0 = heart_disease[heart_disease$Smoking=="No",]
data.1 = heart_disease[heart_disease$Smoking=="Yes",]
b = 0.01
set.seed(2222)
s.0 = data.0[sample(nrow(data.0),size = floor(a*nrow(data.0))),]
s.1 = data.1[sample(nrow(data.1),size = floor(b*nrow(data.1))),]
s.data = rbind(s.0,s.1)
summary(s.data$Smoking)
```

### 3.2 Estimate Propensity Score
```{r}
# estimate propensity score
log.mod = glm(Smoking~.-HeartDisease,data = s.data,family = binomial(link = "logit"))
s.data$ps = predict(log.mod,type = "response")
```

```{r}
histbackback(split(s.data$ps,s.data$Smoking),main="PS before matching",probability = T)
```

### 3.3 Linear regression before matching
```{r}
# Linear regression before matching
l.mod_before = lm(HeartDisease~Smoking+BMI+AlcoholDrinking+Stroke+PhysicalHealth+MentalHealth+DiffWalking+Sex+AgeCategory+Race+Diabetic+PhysicalActivity+GenHealth+SleepTime+Asthma+KidneyDisease+SkinCancer,data = s.data)
summary(l.mod_before)$coefficients[2,]
confint(l.mod_before,parm = 2,0.95)
```

### 3.4 PSM
```{r}
#PSM
match.res2 = matchit(Smoking~BMI+AlcoholDrinking+Stroke+PhysicalHealth+MentalHealth+DiffWalking+Sex+AgeCategory+Race+Diabetic+PhysicalActivity+GenHealth+SleepTime+Asthma+KidneyDisease+SkinCancer,data = s.data,method = "nearest",caliper = 0.05)
matchsum2 = summary(match.res2,standardize = T)$sum.matched
summary(match.res2,standardize = T)$nn
data.frame(round(matchsum2[,1:3],3))
matched_data2 = match.data(match.res2)

histbackback(split(matched_data2$ps,matched_data2$Smoking),main="PS after matching",probability = T)
plot(match.res2,type = "hist")
match.matrix2 = data.frame(match.res2$match.matrix)
T_index2 = match(row.names(match.matrix2),row.names(matched_data2))
C_index2 = match(match.matrix2$match.res2.match.matrix,row.names(matched_data2))
Tgroup2 = matched_data2[T_index2,]
Cgroup2 = matched_data2[C_index2,]
m2.test=t.test(Tgroup2$HeartDisease,Cgroup2$HeartDisease,paired = T)
m2.test
```

### 3.5 Linear regression after matching
```{r}
#LR after matching
l.mod_after = lm(HeartDisease~Smoking+AlcoholDrinking+BMI+Stroke+PhysicalHealth+MentalHealth+DiffWalking+Sex+AgeCategory+Race+Diabetic+PhysicalActivity+GenHealth+SleepTime+Asthma+KidneyDisease+SkinCancer,data = matched_data2)
summary(l.mod_after)$coefficients[2,]
confint(l.mod_after,2,0.95)
```

### 3.6 IPW
```{r}
#IPW
library(WeightIt)
library(cobalt)
library(survey)
balance_before2 = bal.tab(Smoking~BMI+AlcoholDrinking+Stroke+PhysicalHealth+MentalHealth+DiffWalking+Sex+AgeCategory+Race+Diabetic+PhysicalActivity+GenHealth+SleepTime+Asthma+KidneyDisease+SkinCancer,data = s.data,estimand="ATT",m.threshold=0.1)$Balance
balance_before2[,1:3]
weight.res2 = weightit(Smoking~BMI+AlcoholDrinking+Stroke+PhysicalHealth+MentalHealth+DiffWalking+Sex+AgeCategory+Race+Diabetic+PhysicalActivity+GenHealth+SleepTime+Asthma+KidneyDisease+SkinCancer,data = s.data,estimand = "ATT",method = "ps")
balance_after2 = bal.tab(weight.res2,thresholds = c(m=0.1))$Balance
balance_after2[,c(1,3,4)]
d.w2 = svydesign(ids = ~1,weights = get.w(weight.res2),data = s.data)
w.mod = svyglm(HeartDisease ~ Smoking,design = d.w2)
summary(w.mod)$coefficients[2,]
confint(w.mod,2,0.95)
```

### 3.7 Results
```{r}
#Results
smoke.res = data.frame(method = c("LR before","PSM","LR after","IPW"),est_ATT=c(summary(l.mod_before)$coefficients[2,1],m2.test$estimate,summary(l.mod_after)$coefficients[2,1],summary(w.mod)$coefficients[2,1]),CI.lower = c(confint(l.mod_before,2,0.95)[1],m2.test$conf.int[1],confint(l.mod_after,2,0.95)[1],confint(w.mod,2,0.95)[1]),CI.upper= c(confint(l.mod_before,2,0.95)[2],m2.test$conf.int[2],confint(l.mod_after,2,0.95)[2],confint(w.mod,2,0.95)[2]),Pvalue = c(summary(l.mod_before)$coefficients[2,4],m2.test$p.value,summary(l.mod_after)$coefficients[2,4],summary(w.mod)$coefficients[2,4]),significant=c("Yes","Yes","Yes","Yes"))
smoke.res
```








